\documentclass[10pt,letterpaper,draftclsnofoot,onecolumn]{IEEEtran}
    
\usepackage{graphicx}                                        
\usepackage{amssymb}                                         
\usepackage{amsmath}                                         
\usepackage{amsthm}                                          

\usepackage{alltt}                                           
\usepackage{float}
\usepackage{color}
\usepackage{url}

\usepackage{balance}
\usepackage[TABBOTCAP, tight]{subfigure}
\usepackage{enumitem}
%\usepackage{pstricks, pst-node}

\usepackage{listings}
% Taken from http://timmurphy.org/2014/01/27/displaying-code-in-latex-documents/ with some modifications
\lstset{
    frame=none,
    tabsize=4, % tab space width
    showstringspaces=false, % don't mark spaces in strings
    numbers=none, 
    commentstyle=\color{green}, % comment color
    keywordstyle=\color{blue}, % keyword color
    stringstyle=\color{red}, % string color
    basicstyle={\small\ttfamily}, % font style
    belowcaptionskip=1\baselineskip
}

\usepackage{geometry}
\geometry{margin=0.75in}

%random comment

\newcommand{\cred}[1]{{\color{red}#1}}
\newcommand{\cblue}[1]{{\color{blue}#1}}

\newcommand{\toc}{\tableofcontents}

\usepackage{hyperref}

\def\name{Will Sims, Aidan O'Malley}

%pull in the necessary preamble matter for pygments output
%\input{pygments.tex}

%% The following metadata will show up in the PDF properties
\hypersetup{
    colorlinks = false,
    urlcolor = black,
    pdfauthor = {\name},
    pdfkeywords = {cs444},
    pdftitle = {CS 444 Project 4: 4},
    pdfsubject = {CS 444 Project 4},
    pdfpagemode = UseNone
}

\parindent = 0.0 in
\parskip = 0.1 in

\begin{document}

%input the pygmentized output of mt19937ar.c, using a (hopefully) unique name
%this file only exists at compile time. Feel free to change that.

\begin{titlepage}
\title{CS 444 Project 3: Encrypted Block Device}
\author
{\IEEEauthorblockN{Will Sims, Aidan O'Malley\\
}
\IEEEauthorblockA{
CS 444\\
Fall 2017
}}
    \maketitle
    \vspace{2cm}
    \begin{abstract}
        \noindent This document contains the write-up for Project 4 of Operating Systems II. 
        We modified an existing SLOB allocator implementation to use a best-fit algorithm. 

    \end{abstract}

\end{titlepage}

\section{Design for SLOB SLAB}
\noindent{The SLOB allocator is a small and efficient allocator that uses a first-fit algorithm for memory allocation.
Before we started programming we first looked up documentation to understand the slob implementation and best-fit algorithms. 
After viewing slob.c, we noticed that the slob\_alloc() implements a first-fit algorithm to search for a page that has enough space for a new request.
We plan to modify slob\_alloc() and used this pseudocode as a reference for our best-fit algorithm: 

\lstset{language=C++,caption={Best-fit algorithm pseudocode\cite{bestfit}.}}
\begin{lstlisting}
size(block) = n + size(header) 
Scan free list for smallest block with nWords >= size(block) 
if block not found 
    Failure (Garbage collection) 
else if free block nWords >= size(block) + threshold* 
    Split into a free block and an in-use block 
    Free block nWords = Free block nWords - size(block) 
    In-use block nWords = size(block) 
    Return pointer to in-use block 
else 
    Unlink block from free list 
    Return pointer to block 
\end{lstlisting}

Reference: https://www.cs.rit.edu/~ark/lectures/gc/03\_03\_03.html

}

\section{Version Control Log}
\noindent Table from repo logs:

\begin{center}
    \begin{tabular}{ | p{8cm} | p{3cm} | p{6cm} |}
    \hline
    Commit Link & Author & Commit Message \\ \hline
    \href{https://github.com/omalleya/cs444-concurrency-writeups/commit/1c1a0c751d5fe807aa23f82d29d99dda6977e900}{1c1a0c751d5fe807aa23f82d29d99dda6977e900} & William Sims & adds project 4 write-up \\ \hline
    0ab81e85157f5ee29765150f8c76ab9fe850102e & omalleya & adds changes to slob\_alloc to implement best fit and create mem\_usage\ syscall \\ \hline
 
    \end{tabular}
\end{center}

\section{Work Log}
\noindent What work was done and when:
\begin{description}
\item [11/30/17] Met at Kelley at 9:30 and worked on the assignment until 1:30.

\end{description}

\section{Questions}

\noindent\textbf{What do you think the main point of this assignment is?}

\indent{ The main point of this assignment is learn about memory allocation in the Linux kernel.
We will gain experience implementating a best-fit algorithm and how to modify existing existing code. 
We also will learn how to compute the effeciency of the first-fit algorithm and best-fit algorithm and compare the fragmentation sufferred by each algorithm.
We will learn about how to examine the tradeoffs between each of the implmentations. 
}

\noindent\textbf{How did you personally approach the problem?}

\indent{ Our personal apparoch to the problem started with understanding the functions in the slob.c file.
We researched best-fit algorithms for allocating memory and used pseudocode from https://www.cs.rit.edu/~ark/lectures/gc/03\_03\_03.html as the basis for our algorithm design.
In order to implement best fit we need to find the page that has the minimum amount of space free for the required amount that needs to be allocated.

}

\noindent\textbf{How did you ensure your solution was correct?}

\indent{
We ensured out solution was correct by writing a test program that allocates memory and outputs the amount of memory used each time. 
When comparing the first-fit and best-fit slob allocators, we would expect the best-fit allocator to use less memory when the test program runs. 
Best-fit should use less memory because first-fit allocates memory from the fist block it encounters large enough to hold the request while best fit looks for the minimum sized block that can hold the requested amount.

}

\noindent\textbf{What did you learn?}

\indent{ We learned about how memory is allocated in the kernel and how to modify an existing slob allocator implementation to use a best-fit allgorithm. 

}

\noindent\textbf{How should the TA evaluate your work? Provide detailed steps to prove correctness.}
\begin{description}
\item \texttt{[Step: 1] START FROM CLEAN linux-yocto v3.19.2}
\item \texttt{[Step: 2] git apply assign4.patch}
\item \texttt{[Step: 3] source /scratch/opt/poky/1.8/environment-setup-i586-poky-linux}
\item \texttt{[Step: 4] cp /scratch/files/core-image-lsb-sdk-qemux86.ext4 *top of linux tree* \newline //copy driver file if not inside already}
\item \texttt{[Step: 5] cp /scratch/files/config-3.19.2-yocto-standard *top of linux tree* \newline//copy config if not inside already}
\item \texttt{[Step: 6] make menuconfig}
\item \texttt{[Step: 7] General Setup}
\item \texttt{[Step: 8] Change local version name to “-group-19”}
\item \texttt{[Step: 9] From General Setup go to "Choose SLAB allocator" and select SLOB}
\item \texttt{[Step: 10] Save and Exit}
\item \texttt{[Step: 11] make -j4 all \newline //build kernel}
\item \texttt{[Step: 12] command to start VM: \newline qemu-system-i386 -gdb tcp::5519 -S -nographic -kernel arch/x86/boot/bzImage -drive file=core-image-lsb-sdk-qemux86.ext4 -enable-kvm -net nic -net user,hostfwd=tcp::8000-:22 -usb -localtime --no-reboot --append "root=/dev/hda rw console=ttyS0 debug"}
\item \texttt{[Step: 13] \$GDB \newline //From another terminal}
\item \texttt{[Step: 14] target remote :5519}
\item \texttt{[Step: 15] continue}
\item \texttt{[Step: 16] scp -P 8000 ./test.c root@localhost:~ \newline //From another terminal}
\item \texttt{[Step: 17] scp -P 8000 ./test-2.c root@localhost:~}
\item \texttt{[Step: 18] cd ./testmodule && make}
\item \texttt{[Step: 19] scp -P 8000 ./test.ko root@localhost:~ \newline //transfers over module}
\item \texttt{[Step: 20] Login to VM as root and test with the following}
\item \texttt{[Step: 21] compile test.c files with gcc}
\item \texttt{[Step: 22] run to see current memory usage}
\item \texttt{[Step: 23] insmod test.ko \newline //loads module}
\item \texttt{[Step: 24] run test files again to see memory usage}
\item \texttt{[Step: 25] repeat steps 1-24 to test old slob but first rename mm/slob.c to mm/slob-new.c and renaming slob-old.c to slob.c}
\item \texttt{[Step: 26] also run "make -j4 all" to reflect the changes to slob}
\end{description}

\bibliography{hw4}
\bibliographystyle{IEEEtran}

\end{document}
    
